LHC abort gap monitor electronics upgrade

### 2024-06-11 paper.tex

\begin{abstract}
    The LHC Abort Gap Monitor (AGM) is part of the LHC machine protection
    system (MPS) and is designed to measure the particle population in a \SI{3}{\micro\second}
    wide region known as the "abort gap". This region needs to be kept empty to
    ensure safe beam dumps. The AGM captures the synchrotron light generated in
    the visible part of the spectrum and converts it into an electric signal.
    This signal is then processed by an acquisition system and can trigger the
    ‘abort gap cleaning’ process. The current AGM, which has been in operation
    since 2010, uses an analogue integrator ASIC and a 40 MHz
    analogue-to-digital (ADC) converter to provide the particle population
    information. However, this solution is now considered obsolete and is being
    replaced by a digital signal processing approach. Working directly in the
    digital domain not only offers more scalability but also better determinism
    and reliability. This work presents the new technical solution for the
    acquisition chain, compares the characteristics of both implementations,
    and showcases recent measurements conducted on the LHC ion and proton
    beams.
\end{abstract}

\section{Introduction}
The LHC ring stores up to \SI{350}{MJ} of beam energy, when operating at the flat top energy, spread across 2808 bunches with an average beam transverse size under \SI{1}{mm}.
%
This high energy density can melt metallic surfaces and quench
superconducting magnets, posing a risk to the machine~\cite{LHC_report}.
%
To safely extract the beam from the circulating beam lines, a~series of kicker magnets belonging to the beam dump system is used. 
%
Kicker magnets need \SI{3}{\micro\second} to ramp up their magnetic field~\cite{beam_dump_system}.
%
This defines the size of an Abort Gap (AG) in the bunch train.
%
Due to system imperfections, a residual amount of particles populates this gap.
%
The typical AG population levels are two to three orders of magnitude lower in comparison to a single pilot LHC bunch.
%
When the beam dump is initiated, the particles in the AG do not receive full
extraction force, leading to losses along the dump line that could result in damage or magnet quenching.
%
%That causes the particles in the AG to
% deflect towards the vacuum chamber, possibly causing the magnets to quench. 
%
The purpose of the AGM is a~continuous monitoring of the AG population in order to trigger the abort gap cleaning when  necessary \cite{particles_in_ag} .
\\
This paper discusses the upgrade of the operational AGM installed in the LHC since 2010, detailing the system architecture and analogue electronics chain, and provides a first performance comparison  of both systems.

@@

%
The AGM is installed at LHC IP4 as one of the beam diagnostics based on synchrotron radiation (SR) detection. %together with other beam diagnostic, and the LHC low-level RF system.
%
The LHC timing distribution (TTC) broadcasts through an optical cable - along the LHC - turn and bunch marks, together with other information, as e.g., actual beam energy~\cite{ttc_distribution}.
%
The bunch and turn marks identify the start of each individual bunch, and the position of the first bunch in a turn.   
%
The AGM uses this information to re-synchronize its acquisition window to the AG.
%
\subsection{Working principle}
%
An in-vacuum mirror intercepts  SR  emitted by protons passing through magnets located approximately 27 meters upstream. It then directs the radiation to a dedicated optical line installed beneath the beam pipe.

%The particle population in the AG is measured intercepting the synchrotron radiation (SR). 
%
The AGM collects part of the extracted SR via Multi-channel plate photo-multiplier (MCP-PMT) converting it into an electric signal.
%
Neutral density filters in the optical front-end (OF) allow  changing the light intensity, thus optimizing the system dynamic range.
% 
Subsequently, the electric signal is amplified and transmitted over a~\SI{100}{m}-long cable to DAQ.
%
The DAQ integrates the incoming signal in \SI{100}{\nano\second} intervals within a \SI{6.4}{\micro\second}-long acquisition window.
%
Each of the 64 \SI{100}{\nano\second} integrals is independently averaged in an Exponential Moving Averager (EMA).
%
The averaged result is published to the software with a repetition rate of \SI{10}{Hz}, which is sufficient to react to even the fastest changes in AG population.~\cite{high_sensitivity_measurement}.
%
\\
The measurement period depends on the specific MCP-PMT (HAMAMATSU R5916U-50) that features a maximum duty cycle of \SI{1}{\%}. 
%is determined by the PMT used.
%
%In our case, it is HAMAMATSU R5916U-50. 
%
%Its maximum duty cycle is \SI{1}{\%}. 
%
With a gate pulse duration of \SI{3}{\micro\second} and one LHC revolution period of \SI{89}{\micro\second}, the measurements are taken every $4^{\text{th}}$ clock cycle (\SI{356}{\micro\second}).
%
This corresponds to a~duty cycle of \SI{0.84}{\%}.
\\
%

\subsection{Operational system, used since 2010}
The currently used system is shown in Fig.~\ref{fig:current_system_processing}.
%
The optical path, including filters and MCP-PMT, is common to both the operational version of the system, and the proposed upgrade.
\begin{figure}[!tbh]
    \centering
    \includegraphics[width=\columnwidth]{existing_system_processing.pdf}
    \caption{Operational AGM system overview}
    \label{fig:current_system_processing}
\end{figure}
%As well, the PMT type is the same in both systems.
The operational setup utilizes a digital acquisition board (DAB64x) incorporating an Altera Stratix EP1S40 FPGA for data processing.
%
The DAB64x hosts two Individual Bunch Measurement Systems (IBMS)~\cite{ibms}, integrating the analog input signal via the LHCb2002 ASIC~\cite{lhcb_asic}.
%
The analog signal generated by  the MCP-PMT is amplified  prior to integration to  enhance the measurement's Signal-to-Noise Ratio  (SNR).
%
The amplifier (HAMAMATSU C5594~\cite{bsra_first_operation}) has a fixed voltage gain of \SI{36}{dBV} and the bandwidth of \SI{1.5}{GHz} with the low cutoff frequency at \SI{50}{kHz}.
%
The amplified signal is fed to the LHCb2002 in the IBMS integrating in \SI{25}{\nano\second} intervals, producing a stream of analog  'DC-like' values.
%
These are subsequently sampled by a \SI{40}{MSPS} 14-bit ADC, resulting in a digital data stream, where each value corresponds to an integral of an individual LHC \SI{25}{ns} bucket.
%
The DAQ gateware sums four consecutive samples to compute the
proton charge present on a~\SI{100}{ns} interval.
%
This process generates 64 values per single turn acquisition.
%
For each individual value, a standard deviation and mean value is calculated, and separately filtered by EMA. 
%
All collected data are stored in a RAM block shared to a front-end computer (FEC) via VME64 bridge.
%
A dedicated  front-end  server, publishes the data stream through FESA~\cite{fesa}.

@@

The operational AGM primarily relies on outdated   technology, such as  the analogue integrators used in the IBMS,  which are no longer manufactured.
%that are not produced anymore.  manufactured using obsoleted technology, rendering them impossible to get.
%
Furthermore, each integrator is unique, with varying gain, switching characteristics, and offsets. The gateware must correct these variations.
%Furthermore,  each integrator is unique, with varying gain, switching characteristics, and offsets. 
%
%The gateware must correct these variations.
%different gain, different switching characteristics, different offsets and the 
%
%So, the
%gateware must correct these effects.
%
In addition, the amplifier bandwidth (50 kHz-1,5 GHz) %increases system complexity, as
required a correction for the missing DC component and the amplifier gain is fixed,  for which  the overall dynamic range is limtied by the ADC. 
%needs to be implemented.
%
%dynamic range is not fully exploited. cannot optimally re-use available dynamic range.
\\
The system relies on an outdated FPGA platform, and the development SW is no more compatible with the latest operating systems.

\section{System Upgrade}
To address the long-term maintenance concerns, the original system described above has to be refurbished. 
%
While the optics of the new system was copied from the operational one, the DAQ and the front-end amplifier were replaced by a modern solution.
%
A new DC-coupled amplifier was designed, thus allowing enhancing the overall dynamic range. 
%
It was decided to change from analog to digital integration of the MCP-PMT signal, via a DAQ based on  a~4-channel \SI{500}{MSPS} 14-bit FMC-form factor sampler.
%
The FMC ADC module is installed into an in-house built FPGA carrier (VFC-HD), featuring 2$\times 8$GBits DDR3 memories and a large Arria V FPGA.
%
Such a solution eliminates completely the need for the obsolete analogue integrators, as all the signal processing can be implemented in the gateware.
\\
In early 2023, the new AGM system was installed alongside the legacy system to facilitate an easy comparison of both systems' performance.
%An updated version of the AGM was in 2023 installed \textit{in parallel to the actually running AGM}, so that we can compare behavior of both systems.

@@

Figure \ref{fig:amp_schematics} depicts the block diagram of the new amplifier.
%
The amplifier was designed to match the AGM MCP-PMT and to be  highly configurable.
% A = U^2/(i^2 * R^2) => A_dB = 20*log10(U/(i*R))
%
A universal asynchronous receiver-transmitter (UART) interface allows for setting a~trans-impedance gain.
%
The gain is adjustable in a~range of \SI{5.8}{mV/\micro A} to \SI{1.1}{V/\micro A}.
%
This corresponds to a~voltage gain of \SI{41}{dB} to \SI{87}{dB} in a $50\Omega$ system.
%
A programmable DC-offset compensation is implemented in the amplifier.
%
It can be used - together with variable gain - to optimize the signal's amplitude for the ADC.
%
The amplifier design incorporates an alarm. This alarm interlocks the MCP-PMT gate when the incident light produces a current nearing the MCP-PMT's maximum allowable average current.
%The amplifier implements 
%an alarm that interlocks the MCP-PMT gate when the incident light generates a current approachng  the MCP-PMT's maximum allowable average current.
%
The alarm threshold setting is stored through the UART into the amplifier's non-volatile memory, so that 
%it protects 
the MCP-PMT is protected also after a power reset. %immediately at the amplifier power-up.
%
%The alarm status is also transmitted to the gateware.
%
\begin{figure}[!tbh]
    \centering
    \includegraphics[width=\columnwidth]{amp_schematics.pdf}
    \caption{New AGM custom amplifier.}
    \label{fig:amp_schematics}
\end{figure}
%
\subsection{Acquisition system}
The new DAQ (Fig.~\ref{fig:bsra_daq}) receives asynchronous data samples as a stream of four frames every~\SI{8}{\nano s}, each corresponding to one channel.
%
\begin{figure}[!tbh]
    \centering
    \includegraphics[width=\columnwidth]{bsra_daq.pdf}
    \caption{Acquisition system schematics.}
    \label{fig:bsra_daq}
\end{figure}
%
Each frame contains 4$\times$16-bit words, storing left-aligned 14-bit ADC samples. 
%
To compute integrals over 4$\times$~\SI{25}{\nano\second} bunch slots, the system employs data tagging. 
%
The tagger resamples the LHC timing, and generates timing masks that  provide each individual ADC frame with %allowing for identification of 
bunch and turn start markers. 
%within each individual ADC frame.
%
Tags are inserted into the  2 unused least significant bits (LSBs) of each word to indicate the start of the  the bunch or turn.
\\
%
The tagged samples are stored in a~FIFO (First In, First Out), and then integrated. 
%
The integrator is triggered %begins summing values 
by the first bunch tag appearing within the acquisition window.
%
In parallel the system reads a stream of 64~values of a stored 'offset' from a pedestal memory (P[63:0]).
%
The offset is subtracted  from the integrated data stream on the fly.
%
Such pedestal subtraction is needed to remove systematic effects, e.g. related to the  MCP-PMT gate opening. 
%
The output values are filtered by EMA, the results are stored  in RAM, and then read out through the VME64x bus~\cite{my_thesis}.

\begin{figure}[!tbh]
    \centering
    \includegraphics[width=1.0\columnwidth]{installation.jpg}
    \caption{New AGM installed in the LHC IP4, next to BSRT.}
    \label{fig:installation}
\end{figure}

@@

The new installation, when \textit{in retracted position} is  highlighted by a~violet frame in Fig.~\ref{fig:installation}, 
%
The operational AGM  (on the right) and  the longitudinal density monitor (LDM, on the left) are also partially visible in the same figure. 
% Adjacent to its right is the operational AGM, while the longitudinal density monitor (LDM) occupies the left side.
%
All these instruments are part of the LHC beam synchrotron light telescope (BSRT) and use part of the synchrotron radiation extracted by the in-vacuum extraction system sitting above the BSRT.
% instruments share the same light source provided by the LHC beam synchronous light telescope (BSRT).
%
The red arrow indicates the direction of the light entering the optical table.
%
The light is unequally split using multiple reflective foils into their respective destinations.  The new AGM receives  only 10\% of the radiation sent to the operational AGM.
%
Both AGMs are equipped with LEDs installed  on top of their reflective foils for functional and calibration tests. 
%
%These serve to perform functional tests.
%
The new AGM amplifier is mounted on the top of the optical front-end to ease the interventions.

@@

\vspace{-5mm}
\begin{figure}[!htb]
    \begin{center}
        \scalebox{0.54}{\input{plots/ions.pgf}}
        \caption{Comparison of the AGM measurements}
        \label{fig:comparison_chart}
    \end{center}
\end{figure}

\vspace{-5mm}
Initial measurements were conducted in November 2023 using the LHC ion beam.
%
Figure~\ref{fig:comparison_chart} shows the very first measurements, compared to the operational AGM.
%
Both measurements agree to 2\% within the operational area, but deviate starting at $47^{\text{th}}$ bin, resulting in maximum relative error of 4\%.
Despite the MCP-PMT gate is closed at that point and no signal is present, the operational AGM signal gradually decays, and extends below zero.
%
%Whoops, no signal shall be present - the MCP-PMT gates are closed.
%This is surprising, as no signal shall be present - the MCP-PMT gates are closed. 
%
This effect is systematic, and is mostly caused by analog signal processing chain and non-ideal analogue integration.
%
The new system effectively eliminates this.

\section{Summary}
This paper describes upgrade of the operational AGM.
%
The operational system has to be replaced in the near future by a new AGM to assure long-term maintainability.
%
A fully digital solution was developed.
%
To assess its performance, it was installed in the LHC, on beam 1, in parallel with the operational AGM.
%
Very first measurements from November 2023 show exceptionally good performance even using considerably lower amount of intercepted light.
%
The upgraded AGM is now collecting data from the LHC 2024 proton run.
%
Further analysis in terms of accuracy, reliability, and stability has to be conducted, with the ultimate goal of replacing both operational AGMs by their new versions.

>>>>

% NOTE: I (David) had to remove from abstract the proton beam info, because we did not
% have enough time to work with protons. This article only compares ION beams so far.
% I'll try to tweak the submitted abstract as well - if that is still possible. If not,
% the correct abstract will be shown in the article.
\begin{abstract}
    The LHC Abort Gap Monitor (AGM) is part of the LHC machine protection
    system (MPS) and is designed to measure the particle population in a \SI{3}{\micro\second}
    wide region known as the "abort gap". This region needs to be kept empty to
    ensure safe beam dumps. The AGM captures the synchrotron light generated in
    the visible part of the spectrum and converts it into an electric signal.
    This signal is then processed by an acquisition system and can trigger the
    ‘abort gap cleaning’ process. The current AGM, which has been in operation
    since 2010, uses an analogue integrator ASIC and a 40 MHz
    analogue-to-digital (ADC) converter to provide the particle population
    information. However, this solution is now considered obsolete and is being
    replaced by a digital signal processing approach. Working directly in the
    digital domain not only offers more scalability but also better determinism
    and reliability. This work presents the new technical solution for the
    acquisition chain, compares the characteristics of both implementations,
    and showcases recent measurements conducted on the LHC ion beams.
\end{abstract}

\section{Introduction}
The LHC ring stores up to \SI{350}{MJ} of beam energy, when operating at the flat top energy, spread across 2808 bunches with an average beam transverse size under \SI{1}{mm}.
%
This high energy density can melt metallic surfaces and quench
superconducting magnets, posing a risk to the machine~\cite{LHC_report}.
%
To safely extract the beam from the circulating beam lines, a~series of kicker magnets belonging to the beam dump system is used. 
%
Kicker magnets need \SI{3}{\micro\second} to ramp up their magnetic field~\cite{beam_dump_system}.
%
This defines the duration of an Abort Gap (AG) in the bunch train.
%
Due to system imperfections, a residual amount of particles populates this gap.
%
The typical AG population levels are two to three orders of magnitude lower in comparison to a single pilot LHC bunch.
%
When the beam dump is initiated, the particles in the AG do not receive full
extraction kicks, leading to losses along the dump line that could result in damage or magnet quenching.
%
%That causes the particles in the AG to
% deflect towards the vacuum chamber, possibly causing the magnets to quench. 
%
The purpose of the AGM is a~continuous monitoring of the AG population in order to trigger the abort gap cleaning when  necessary \cite{particles_in_ag}.
\\
This paper discusses the upgrade of the operational AGM installed in the LHC since 2010, detailing the system architecture and analogue electronics chain, and provides a first performance comparison  of both systems.

@@

%
The AGM is installed at LHC IP4 as one of the beam diagnostics based on synchrotron radiation (SR) detection. %together with other beam diagnostic, and the LHC low-level RF system.
%
The LHC timing distribution (TTC) broadcasts along the LHC turn and bunch marks, together with other information, as e.g., actual beam energy~\cite{ttc_distribution}.
%
The bunch and turn marks identify the start of each individual bunch, and the position of the first bunch in a turn.   
%
The AGM uses this information to re-synchronize its acquisition window to the AG.
%
\subsection{Working principle}
%
An in-vacuum mirror intercepts  SR  emitted by particles passing through magnets located approximately \SI{27}{m} upstream. It then directs the radiation to a dedicated optical line installed beneath the beam pipe.

%The particle population in the AG is measured intercepting the synchrotron radiation (SR). 
%
The AGM collects part of the extracted SR via Multi-channel plate photo-multiplier (MCP-PMT) converting it into an electric signal.
%
Neutral density filters in the optical front-end (OF) allow  changing the light intensity, thus optimize the system dynamic range.
% 
Subsequently, the electric signal is amplified and transmitted over a~\SI{100}{m}-long cable to DAQ.
%
The DAQ integrates the incoming signal in \SI{100}{\nano\second} intervals within a \SI{6.4}{\micro\second}-long acquisition window.
%
Each of the 64 \SI{100}{\nano\second} integrals is independently averaged in an Exponential Moving Averager (EMA).
%
The averaged result is published to the software with a repetition rate of \SI{10}{Hz}, which is sufficient to react to even the fastest changes in AG population.~\cite{high_sensitivity_measurement}.
%
\\
The measurement period depends on the specific MCP-PMT (HAMAMATSU R5916U-50) that features a maximum duty cycle of \SI{1}{\%}. 
%is determined by the PMT used.
%
%In our case, it is HAMAMATSU R5916U-50. 
%
%Its maximum duty cycle is \SI{1}{\%}. 
%
With a gate pulse duration of \SI{3}{\micro\second} and one LHC revolution period of \SI{89}{\micro\second}, the measurements are taken every $4^{\text{th}}$ clock cycle (\SI{356}{\micro\second}).
%
This corresponds to a~duty cycle of \SI{0.84}{\%}.
\\
%

\subsection{Operational system, used since 2010}
The currently used system is shown in Fig.~\ref{fig:current_system_processing}.
%
The optical path, including filters and MCP-PMT, is common to both the operational version of the system, and the proposed upgrade.
\begin{figure}[!tbh]
    \centering
    \includegraphics[width=\columnwidth]{existing_system_processing.pdf}
    \caption{Operational AGM system overview}
    \label{fig:current_system_processing}
\end{figure}
%As well, the PMT type is the same in both systems.
The operational setup utilizes a digital acquisition board (DAB64x) incorporating an Altera Stratix EP1S40 FPGA for data processing.
%
The DAB64x hosts two Individual Bunch Measurement Systems (IBMS)~\cite{ibms}, integrating the analog input signal via the LHCb2002 ASIC~\cite{lhcb_asic}.
%
The analog signal generated by  the MCP-PMT is amplified  prior to integration to  enhance the measurement's Signal-to-Noise Ratio  (SNR).
%
The amplifier (HAMAMATSU C5594~\cite{bsra_first_operation}) has a fixed voltage gain of \SI{36}{dBV} and the bandwidth of \SI{1.5}{GHz} with the low cutoff frequency at \SI{50}{kHz}.
%
The amplified signal is fed to the LHCb2002 in the IBMS integrating in \SI{25}{\nano\second} intervals, producing a stream of analog  'DC-like' values.
%
These are subsequently sampled by a \SI{40}{MSPS} 14-bit ADC, resulting in a digital data stream, where each value corresponds to an integral of an individual LHC \SI{25}{ns} bucket.
%
The DAQ gateware sums four consecutive samples to compute the charge present on a~\SI{100}{ns} interval.
%
This process generates 64 values per single turn acquisition.
%
For each individual value, a standard deviation and mean value is calculated, and separately filtered by EMA. 
%
All collected data are stored in a RAM block shared to a front-end computer (FEC) via VME64 bridge.
%
A dedicated  front-end  server publishes the data stream through FESA~\cite{fesa}.

@@

The operational AGM primarily relies on outdated   technology, such as  the analogue integrators used in the IBMS,  which are no longer manufactured.
%that are not produced anymore.  manufactured using obsoleted technology, rendering them impossible to get.
%
Furthermore, each integrator is unique, with varying gain, switching characteristics, and offsets. The gateware must correct these variations.
%Furthermore,  each integrator is unique, with varying gain, switching characteristics, and offsets. 
%
%The gateware must correct these variations.
%different gain, different switching characteristics, different offsets and the 
%
%So, the
%gateware must correct these effects.
%
In addition, the amplifier bandwidth (50 kHz-1,5 GHz) %increases system complexity, as
requires a correction for the missing DC component.
%%%%%%%% and the amplifier gain is fixed,  for which  the overall dynamic range is limited by the ADC. 
% To Federico: The statement above is not true. ADC dynamic range is determined solely by number of bits. 
% What changes is the SNR, which is defined as power in signal amplitude wrt to power of noise. 
% In ideal conditions we want our signal to be amplified such, that we use maximum available ADC 
% analogue signal amplitude to increase signal to noise ratio. ADC dynamic range is constant.
% I'd replace it by following:
The amplifier gain is fixed,  so the signal amplitude cannot be adapted to optimally re-use the available ADC dynamic range. 
% this is what happens electrically.
\\
The system also relies on an outdated FPGA platform, and the development SW is no more compatible with the latest operating systems.

\section{System Upgrade}
To address the long-term maintenance concerns, the original system described above has to be refurbished. 
%
While the optics of the new system was copied from the operational one, the DAQ and the front-end amplifier were replaced by a modern solution.
%
A new DC-coupled variable gain amplifier was designed.
%
It was decided to change from analog to digital integration of the MCP-PMT signal.
%
The signal is digitized by a DAQ based on a~4-channel \SI{500}{MSPS} 14-bit FMC-form factor sampler.
%
The FMC ADC module is installed into an in-house built FPGA carrier (VFC-HD), featuring 2$\times 8$GBits DDR3 memories and a large Arria V FPGA.
%
Such a solution eliminates completely the need for the analogue integrators, as all the signal processing can be implemented in the gateware.
\\
In early 2023, the new AGM system was installed alongside the legacy system to compare both systems' performance.
%An updated version of the AGM was in 2023 installed \textit{in parallel to the actually running AGM}, so that we can compare behavior of both systems.

@@

Figure \ref{fig:amp_schematics} depicts the block diagram of the new amplifier.
%
The amplifier was designed to match the AGM MCP-PMT and to be  highly configurable.
% A = U^2/(i^2 * R^2) => A_dB = 20*log10(U/(i*R))
%
A universal asynchronous receiver-transmitter (UART) interface allows for setting a~trans-impedance gain.
%
The gain is adjustable in a~range of \SI{5.8}{mV/\micro A} to \SI{1.1}{V/\micro A}.
%
This corresponds to a~voltage gain of \SI{41}{dB} to \SI{87}{dB} in a $50\Omega$ system.
%
A programmable DC-offset compensation is implemented.
%
It can be used - together with variable gain - to optimize the signal's amplitude for the ADC.
%
The amplifier design incorporates an alarm. This alarm interlocks the MCP-PMT gate when the incident light produces a current nearing the MCP-PMT's maximum allowable average current.
%The amplifier implements 
%an alarm that interlocks the MCP-PMT gate when the incident light generates a current approachng  the MCP-PMT's maximum allowable average current.
%
The alarm threshold setting is stored through the UART into the amplifier's non-volatile memory, so that 
%it protects 
the MCP-PMT is protected also after a power reset. %immediately at the amplifier power-up.
%
%The alarm status is also transmitted to the gateware.
%
\begin{figure}[!tbh]
    \centering
    \includegraphics[width=\columnwidth]{amp_schematics.pdf}
    \caption{New AGM custom amplifier.}
    \label{fig:amp_schematics}
\end{figure}
%
\subsection{Acquisition system}
The new DAQ (Fig.~\ref{fig:bsra_daq}) receives asynchronous data samples as a stream of four frames every~\SI{8}{\nano s}, each corresponding to one channel.
%
\begin{figure}[!t]
    \centering
    \includegraphics[width=\columnwidth]{bsra_daq.pdf}
    \caption{Acquisition system schematics.}
    \label{fig:bsra_daq}
\end{figure}
%
Each frame contains 4$\times$16-bit words, storing left-aligned 14-bit ADC samples. 
%
To compute integrals over 4$\times$~\SI{25}{\nano\second} bunch slots, the system employs data tagging. 
%
The tagger resamples the LHC timing, and generates timing masks that  provide each individual ADC frame with %allowing for identification of 
bunch and turn start markers. 
%within each individual ADC frame.
%
Tags are inserted into the  2 unused least significant bits (LSBs) of each word to indicate the start of the  the bunch or turn.
\\
%
% To Federico: I would personally remove definitions of UART and FIFO, as they are a 'common knowledge'. It would
% save some space. We need to gain somewhere a half of paragraph to fit into 3 pages + refs.
The tagged samples are stored in a~FIFO (First In, First Out), and then integrated. 
%
The integrator is triggered %begins summing values 
by the first bunch tag appearing within the acquisition window.
%
In parallel, the system reads a stream of 64~values of a stored 'offset' from a pedestal memory (P[63:0]).
%
The offset is subtracted  from the integrated data stream on the fly.
%
Such pedestal subtraction is needed to remove systematic effects, e.g. related to the MCP-PMT gate opening. 
%
The output values are filtered by EMA, the results are stored  in RAM, and then read out through the VME64x bus~\cite{my_thesis}.

\begin{figure}[!tbh]
    \centering
    \includegraphics[width=1.0\columnwidth]{installation.jpg}
    \caption{New AGM installed in the LHC IP4, next to BSRT.}
    \label{fig:installation}
\end{figure}
\vspace{-5mm}

@@

The new installation, shown \textit{in retracted position}, is  highlighted by a~violet frame in Fig.~\ref{fig:installation}, 
%
The operational AGM  (on the right) and  the longitudinal density monitor (LDM, on the left) are also partially visible in the same figure. 
% Adjacent to its right is the operational AGM, while the longitudinal density monitor (LDM) occupies the left side.
%
All these instruments are part of the LHC beam synchrotron light telescope (BSRT) and use part of the synchrotron radiation extracted by the in-vacuum extraction system sitting above the BSRT.
% instruments share the same light source provided by the LHC beam synchronous light telescope (BSRT).
%
The red arrow indicates the direction of the light entering the optical table.
%
The light is unequally split using multiple reflective foils into their respective destinations.  The new AGM receives  only 10\% of the radiation sent to the operational AGM.
%
Both AGMs are equipped with LEDs installed  on top of their reflective foils for functional and calibration tests. 
%
%These serve to perform functional tests.
%
The new AGM amplifier is mounted on the top of the optical front-end to ease the interventions.

@@

\vspace{-5mm}
\begin{figure}[!htb]
    \begin{center}
        \scalebox{0.54}{\input{plots/ions.pgf}}
        \caption{Example of a comparison between old and new  AGM measurements, binned in 100\,ns slots (\emph{output filter indices}).  The MCP-PMT gate is \emph{open} between index 17 and 47.}
        \label{fig:comparison_chart}
    \end{center}
\end{figure}

\vspace{-5mm}
A first set of measurements with the new AGM system were taken during the 2023 Ion Run, 
%Initial measurements were conducted in November 2023 using the LHC ion beam.
%
and the comparison to the operational AGM is shown in Figure~\ref{fig:comparison_chart}.
%shows the very first measurements, compared to the operational AGM.
%
During the period in which the MCP-PMT gate is open (between index 17 and 47 in the plot) the agreement between old and new system is within $\pm 2\%$.  
%Both measurements agree to 2\% within the operational area, but deviate starting at $47^{\text{th}}$ bin, resulting in maximum relative error of 4\%.
When the MCP-PMT gate closes at index 47 the operational system exhibits a slow (several \SI{100}{ns} slots) decay to non-physical negative values. 
%Despite the MCP-PMT gate is closed at that point and no signal is present, the operational AGM signal gradually decays, and extends below zero.
%
%Whoops, no signal shall be present - the MCP-PMT gates are closed.
%This is surprising, as no signal shall be present - the MCP-PMT gates are closed. 
%
This effect was found to be  systematic, and traced back to  non-optimal analog signal processing and integration. 
%chain and non-ideal analogue integration.
%
The new system eliminates this effect.

\section{Summary}

The current LHC AGM system requires refurbishing to assure long-term maintainability of the DAQ. This paper outlines the new DAQ design, which is a fully digital solution incorporating digital integration. An initial prototype was installed to monitor the LHC on beam 1, working in tandem with the operational AGM.

In 2023, a series of measurements with ions demonstrated exceptional performance, even with significantly less intercepted light, and showed strong alignment with the legacy system.

The unit is now gathering data with protons, which will facilitate further analysis regarding its accuracy, reliability, and stability. Once the system is fully validated, both LHC AGMs will transition to the new version.

%This paper describes upgrade of the operational AGM.
%
%The operational system has to be replaced in the near future by a new AGM to assure long-term maintainability.
%
%A fully digital solution was developed.
%
%To assess its performance, it was installed in the LHC, on beam 1, in parallel with the operational AGM.
%
%Very first measurements from November 2023 show exceptionally good performance even using considerably lower amount of intercepted light.
%
%The upgraded AGM is now collecting data from the LHC 2024 proton run.
%
%Further analysis in terms of accuracy, reliability, and stability has to be conducted, with the ultimate goal of replacing both operational AGMs by their new versions.
